### 2022-07-16 conclusion/conclusion.tex
geoff ch7-refs

After its first victory, Zipperposition was successfully integrated into
Sledgehammer and it is now part of its default installation. We also noticed
that problems which only Zipperposition can solve do pop up from time to time in
practice. Antoine Defourné also integrated Zipperposition in the TLA+ proof
assistant and observed that it helps prove some problems that were previously
out of reach \cite{ad-21-ho-tlaplus}.

@@

This extension of Ehoh to \ehohii{} finished the third cycle. Inspired by our
approach of extending techniques designed for weaker logics to work in the
context of richer logics, we decided to look for ways in which superposition can
assimilate the most successful techniques of SAT solving. Extension of hidden
literals was straightforward, but finding the right way to tame their infinite
nature in the first-order case, as well as the right way to integrate equality
in their definition, was challenging. Predicate elimination posed challenges in
terms of finding the exact conditions under which it can be integrated with the
saturation loop. We first observed that blocked clause elimination destroys the
completeness of superposition calculus when we noticed that Zipperposition does
not prove some problems when the technique is enabled. After sifting through tens of
benchmarks and carefully examining the debugging information, we realized that
this was not due to a bug in the implementation but due to the incompatibility of
blocked clause elimination with the redundancy criterion. In the end, we were left
somewhat disappointed to learn that SAT techniques do not scale good enough to be used
as superposition simplification techniques. However, we learned that they are
very successful as preprocessing techniques.

@@

\ourpara{Higher-Order Unification} The complete variant of the unification procedure
described in Chapter \ref{ch:unif} is proved complete with substantial
restrictions on the order and kind of applied rules. To further
remove redundancy, we plan to find alternative, less explosive unification rules
that do not compromise completeness. Furthermore, we want to investigate if
there are more combinations of unification rules that lead to redundant unifiers
and find ways to remove those redundancy-inducing combinations.

@@

Lastly, we fixed the order in which Zipperposition applies different unification
rules. It would be interesting to see if applying imitation before projection
(or vice versa) yields performance improvements on some kinds of problems. 
We also postponed the application of the most explosive rules such as iteration
to the very end. However, it would not be surprising that there are many hard hand-crafted
problems on which applying these rules early on would lead to a proof more quickly.

>>>>

After its first victory, Zipperposition was successfully integrated into
Sledgehammer and it is now part of Sledgehammer's default installation. We also noticed
that problems which only Zipperposition can solve do pop up from time to time in
practice. Antoine Defourné also integrated Zipperposition in the TLA+ proof
assistant and observed that it helps prove some problems that were previously
out of reach \cite{ad-21-ho-tlaplus}.

@@

This extension of Ehoh to \ehohii{} finished the third cycle. Inspired by our
approach of extending techniques designed for weaker logics to work in the
context of richer logics, we decided to look for ways in which superposition can
assimilate the most successful techniques of SAT solving. Extension of hidden
literals was straightforward, but finding the right way to tame their infinite
nature in the first-order case, as well as the right way to integrate equality
in their definition, was challenging. Predicate elimination posed challenges in
terms of finding the exact conditions under which it can be integrated with the
saturation loop. We first observed that blocked clause elimination destroys the
completeness of superposition calculus when we noticed that Zipperposition does
not prove some problems when the technique is enabled. After sifting through tens of
benchmarks and carefully examining the debugging information, we realized that
this was not due to a bug in the implementation but due to the incompatibility of
blocked clause elimination with the redundancy criterion. In the end, we were left
somewhat disappointed to learn that SAT techniques do not scale well enough to be used
as superposition simplification techniques. However, we learned that they are
very successful as preprocessing techniques.

@@

\ourpara{Higher-Order Unification} The complete variant of the unification procedure
described in Chapter \ref{ch:unif} is proved complete with substantial
restrictions on the order and kind of applied rules. To further
remove redundancy, we hipe to find alternative, less explosive unification rules
that do not compromise completeness. Furthermore, we want to investigate if
there are more combinations of unification rules that lead to redundant unifiers
and find ways to remove those redundancy-inducing combinations.

@@

Lastly, we fixed the order in which Zipperposition applies different unification
rules. It would be interesting to see if applying imitation before projection
(or vice versa) yields performance improvements on some kinds of problems. 
We also postponed the application of the most explosive rules such as iteration
to the very end. However, it would not be surprising if there are many hard hand-crafted
problems on which applying these rules early on would lead to a proof more quickly.

### 2022-07-16 satfol/satfol.tex
geoff ch7-refs

\begin{abstract}
Optimized SAT solvers not only simplify the clause set during preprocessing, but
they also simplify it during solving. This interleaving of
simplification and solving is called inprocessing. Some preprocessing
techniques have been generalized to first-order logic with equality. In this
\paper, we port inprocessing techniques to work with superposition, a leading
first-order proof calculus, and we strengthen known preprocessing techniques.
Specifically, we look into elimination of hidden literals, variables
(predicates), and blocked clauses. Our evaluation using the Zipperposition
prover confirms that the new techniques usefully supplement the existing
superposition machinery.
\end{abstract}

@@

Automated reasoning tools have become much more powerful in the last few decades
thanks to procedures such as conflict-driven clause learning (CDCL)
\cite{MSLM09HBSAT} for propositional logic and superposition for first-order
logic with equality. However, the effectiveness of these procedures crucially
depends on how the input problem is represented as a clause set. The clause set
can be optimized beforehand (\emph{preprocessing}) or during the execution of
the procedure (\emph{inprocessing}). In this \paper, we lift several
preprocessing and inprocessing techniques from propositional logic to clausal
first-order logic and demonstrate their usefulness in a superposition prover.

@@

We assume the natural extensions of domain, valuation, interpretation, and model
(as defined by Fitting \cite{mf-1996-fol}) from unsorted to many-sorted logic.
We also assume the notion of normal models from Sect.~\ref{sec:pre:fol} (with
corresponding notation) and extend it with the notion of \emph{canonical
models}: A canonical model $\ourmodel$ is a normal model such that for every
element $d$ in one of $\ourmodel$'s domains there exists a ground term $t$ such
that $\ourmodel$ interprets $t$ as $d$. 
We also make use of Herbrand models and Herbrand's theorem \cite[Sect.~5.4]{mf-1996-fol}.
In Herbrand's models the domain consists of all ground terms. These models have a 
useful property that if a formula is satisfiable in a model, then it is also satisfiable in an
Herbrand model. Herbrand's theorem states that every unsatisfiable clause set has (propositionally)
unsatisfiable ground instance.     
Canonical models generalize Herbrand models to first-order logic with equality.

@@

As an optimization, predicate elimination keeps track only of symbols that appear
at most $K_\mathrm{occ}$~times in the clause set. For inprocessing, we use
signals that the prover emits whenever a clause is added to or removed from
the proof state and update the records. At the beginning of the
1st, $(K_\mathrm{iter} + 1)$st, $(2 K_\mathrm{iter} + 1)$st, $\dots{}$ iteration of the given clause procedure's
loop body, predicate elimination is systematically applied to the entire proof
state. The first application of inprocessing amounts to preprocessing. By
default, $K_\mathrm{occ} = 512$ and $K_\mathrm{iter}=10$. The same ideas and
limits apply for blocked clause elimination.

@@

The invalidity of set $E$ from step \ref{step:encoding} is checked using a SAT
solver, which is already integrated in Zipperposition. As modern theorem provers
(including E and Vampire) also use SAT solvers, the method can easily be
implemented.

@@

\ourpara{Satisfiability by Blocked Clause Elimination}
%
\looseness=-1%
Kiesl et al.\ found that blocked clause elimination is especially effective on
satisfiable problems. To corroborate their results and ascertain
whether a combination of predicate elimination and blocked clause elimination
increases the success rate, we evaluate BCE on all 2273
satisfiable or TPTP FOF and CNF problems.
The hardware and CPU time limits are the same as in the experiments above.
Figure~\ref{fig:bce-sat} presents the results.

>>>>

\begin{abstract}
Optimized SAT solvers simplify the clause set not only during preprocessing, but
they also simplify it during solving. This interleaving of
simplification and solving is called inprocessing. Some preprocessing
techniques have been generalized to first-order logic with equality. In this
\paper, we port inprocessing techniques to work with superposition, a leading
first-order proof calculus, and strengthen known preprocessing techniques.
Specifically, we look into elimination of hidden literals, variables
(predicates), and blocked clauses. Our evaluation using the Zipperposition
prover confirms that the new techniques usefully supplement the existing
superposition machinery.
\end{abstract}

@@

Automated reasoning tools have become much more powerful in the last few decades,
thanks to procedures such as conflict-driven clause learning (CDCL)
\cite{MSLM09HBSAT} for propositional logic and superposition for first-order
logic with equality. However, the effectiveness of these procedures crucially
depends on how the input problem is represented as a clause set. The clause set
can be optimized beforehand (\emph{preprocessing}) or during the execution of
the procedure (\emph{inprocessing}). In this \paper, we lift several
preprocessing and inprocessing techniques from propositional logic to clausal
first-order logic, and demonstrate their usefulness in a superposition prover.

@@

We assume the natural extensions of domain, valuation, interpretation, and model
(as defined by Fitting \cite{mf-1996-fol}) from unsorted to many-sorted logic.
We also assume the notion of normal models from Sect.~\ref{sec:pre:fol} (with
corresponding notation) and extend it with the notion of \emph{canonical
models}: A canonical model $\ourmodel$ is a normal model such that for every
element $d$ in one of $\ourmodel$'s domains there exists a ground term $t$ such
that $\ourmodel$ interprets $t$ as $d$. 
We also make use of Herbrand models and Herbrand's theorem \cite[Sect.~5.4]{mf-1996-fol}.
In Herbrand models the domain consists of all ground terms. These models have a 
useful property that a formula is satisfiable if and only if it is satisfiable in an
Herbrand model. Herbrand's theorem states that every unsatisfiable clause set has (propositionally)
unsatisfiable ground instance.     
Canonical models generalize Herbrand models to first-order logic with equality.

@@

As an optimization, predicate elimination keeps track only of symbols that appear
at most $K_\mathrm{occ}$~times in the clause set. For inprocessing, we use
signals that the prover emits whenever a clause is added to or removed from
the proof state and update the records. At the beginning of the
1st, $(K_\mathrm{iter} + 1)$st, $(2 K_\mathrm{iter} + 1)$st, $\dots{}$ iteration of the given clause procedure's
loop body, predicate elimination is systematically applied to the entire proof
state. The first application of inprocessing amounts to preprocessing. After some informal experiments,
we chose $K_\mathrm{occ} = 512$ and $K_\mathrm{iter}=10$ as default values. The analogous optimization and
limits apply for blocked clause elimination.

@@

The invalidity of set $E$ from step \ref{step:encoding} is checked using a SAT
solver, which is already integrated in Zipperposition. As modern theorem provers
(including E and Vampire) also use SAT solvers, this definition set recognition method can easily be
implemented in those provers as well.

@@

\ourpara{Satisfiability by Blocked Clause Elimination}
%
\looseness=-1%
Kiesl et al.\ found that blocked clause elimination is especially effective on
satisfiable problems. To corroborate their results and ascertain
whether a combination of predicate elimination and blocked clause elimination
increases the success rate, we evaluate BCE on all 2273
satisfiable TPTP FOF and CNF problems.
The hardware and CPU time limits are the same as in the experiments above.
Figure~\ref{fig:bce-sat} presents the results.
